%!TEX root = thesis.tex
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\thispagestyle{plain}

% The abstract is a short summary of the thesis. It announces in
% a brief and concise way the scientific goals, methods, and most important
% results. The chapter ``conclusions'' is not equivalent to the abstract!
% Nevertheless, the abstract may contain concluding remarks. The abstract
% should not be discursive. Hence, it cannot summarize all aspects of the thesis
% in very detail. Nothing should appear in an abstract that is not also
% covered in the body of the thesis itself. Hence, the abstract should be the
% last part of the thesis to be compiled by the author.

% A good abstract has the following properties: \emph{Comprehensive:} All major
% parts of the main text must also appear in the abstract. \emph{Precise:}
% Results, interpretations, and opinions must not differ from the ones in the main
% text. Avoid even subtle shifts in emphasis. \emph{Objective:} It may contain
% evaluative components, but it must not seem judgemental, even if the thesis
% topic raises controversial issues. \emph{Concise:} It should only contain the
% most important results. It should not exceed 300--500 words or about one page.
% \emph{Intelligible:} It should only contain widely-used terms. It should
% not contain equations and citations. Try to avoid symbols and acronyms (or at
% least explain them). \emph{Informative:} The reader should be able to quickly
% evaluate, whether or not the thesis is relevant for his/her work.

% An Example: The objective was to determine whether \dots (\emph{question/goal}).
% For this purpose, \dots was \dots (\emph{methodology}). It was found that \dots
% (\emph{results}). The results demonstrate that \dots (\emph{answer}).

Global glacier models are necessary to project future glacier ice mass loss and the resulting consequences, such as sea level rise. Glacier change is controlled by  mass balance, the resulting dynamical geometry adjustments and the interaction between them.
The scarcity of ice volume observations makes the calibration, and especially validation, of glacier evolution models difficult. As a consequence, most global models to date used \vas{} relations to parameterize dynamical adjustment processes. Over the last years, the increasing accuracy and coverage of global data inventories, in combination with new and automated methods as well as increased computing power allowed for the development of more sophisticated models. This study compares the dynamic response of the global glacier model based on scaling relations by \citet{Marzeion2012b} to the globally applicable flowline model based on the Shallow Ice Approximation by \citet{Maussion2019} under controlled boundary conditions.

The re-implemented \vas{} model is able to correctly simulate glacier changes in response to various climatic forcings. The computed glacier changes are conceptually correct and qualitatively comparable to---but generally smaller than---those of the flowline model. The differences between \vas{} model and flowline model are most pronounced for idealized equilibrium experiments and can be attributed to a missing mass-balance-elevation feedback and the overestimation of model-internal time scales. Additionally, these too long time scales result in oscillating adjustments of glacier geometries to step changes in climate.

The results of ice volume change projections for Central Europe and High Mountain Asia over the 21st century forced by CMIP6 data are also comparable to the literature, even though on the conservative side. While the values of ice volume change relative to 2020 may differ substantially between the \vas{} model and the flowline model, the values of absolute ice mass loss are in closer agreement. For example, the projected ice loss relative to 2020 for South Asia East (RGI region 15) ranges between $-41\SI{\pm4}{\percent} ~(0.93\SI{\pm0.10}{\milli\meter\sealevelrise})$ under SSP1-2.6 and $-64\SI{\pm5}{\percent} ~(1.44\SI{\pm0.10}{\milli\meter\sealevelrise})$ under SSP5-8.5 for the \vas{} model, and $-55\SI{\pm5}{\percent} ~(1.00\SI{\pm0.10}{\milli\meter\sealevelrise})$ under SSP1-2.6 and $-76\SI{\pm5}{\percent} ~(1.40\SI{\pm0.09}{\milli\meter\sealevelrise})$ under SSP5-8.5 for the flowline model. The values of both models are generally more comparable for High Mountain Asia (given the larger population size), while estimations of both models for Central Europe the can be outside each others uncertainty bounds (especially under lower warming scenarios).

These results show that the use of ice dynamics is advisable for more realistic glacier change projections. Nevertheless, this study also reinforces the \vas{} model's suitability for certain applications, building on its unique advantages: its computational efficiency, relative simplicity, and capacity to estimate past glacier change from today’s geometries. The implementation of the \vas{} model in the open-source OGGM framework realized in this thesis will allow further model applications and comparison studies in the future.
