%!TEX root = thesis.tex
\chapter{Conclusions}\label{chap:conclusions}
\thispagestyle{plain}

% This chapter contains consequences that derive from your results. It may also
% contain speculations. It may provide suggestions for future studies. Hence,
% the conclusions may provide an outlook and list open questions. Sometimes this
% chapter is part of the discussion. In such a case, the chapter reads
% "Discussion and Conclusions".

Glaciers are controlled by their mass balance and the resulting ice flow. Hence, a glacier model has to incorporate a mass balance model, to account for accumulation and ablation, and a dynamic model, to update glacier geometries in response to mass balance changes. A dynamic model based on scaling relations, such as the here presented combination of volume/area, volume/length and response time scaling, can produce qualitatively comparable results to a state-of-the-art flowline model. Thereby, the results of real world applications like projections of regional glacial evolution forced by general circulation model data are mostly within the uncertainty bounds of each other, while idealized equilibrium experiments expose the limitations and shortcomings of the \vas{} model. The projected mass loss over the 21st century by the \vas{} model (e.g., $0.21\SI{\pm0.01}{\milli\meter\sealevelrise}$ and $4.51\SI{\pm0.46}{\milli\meter\sealevelrise}$ under SSP5-8.5 for RGI region 11 Central Europe  and RGI region 13 Central Asia, respectively) is comparable to the results of the flowline model and other recently published studies, even though the estimates are on the conservative side (especially for SSPs with lower warming scenarios).

The \vas{} model is able to correctly simulate advances and retreats in response to step changes in climate as well as to natural climate variability, it has the ability to reach new equilibriums, and it shows characteristics of a low-pass filter. Due to it's simplicity this is done with a high computational effieciency (less than half of the run time including preprocessing compared to the flowline model). However, changes in ice volume, surface area and glacier length are generally underestimated when compared to the flowline model. The model-internal representation of time scales has been identified as the main source of those differences and other unexpected and unwanted behaviors (such as an oscillating response to step changes in climate). The conducted sensitivity analysis, backed by \citet{Roe2020}, concludes that the used time scale parameterization results in too high values. Hence, the timescales could be exploited as future tuning parameter. Using the geometric timescales based on the terminus ablation rate \citep{Johannesson1989} rather than the mass turnover may improve the results.

Additional errors are introduced by missing physically relevant processes, which are not resolved by the \vas{} model. Those include but are not limited to, a missing mass-balance-elevation feedback, a constant average slope, and no awareness of the surrounding topography. While these processes may be secondary on a global scale, estimates for regions with many small glaciers like the Alps vary significantly between scaling and flowline model. The usage of more complex models is advisable for future projections of glacier mass changes, especially given the nowadays available computational power.

This study detailed the underlying principles and implementation process of a simple global glacier model based on glacier scaling relations \citep{Marzeion2012b} in the OGGM framework. The implementation of the \vas{} model in the OGGM framework is a success on its own. As suggested by \citet{Maussion2019}, the implementation of simpler approaches, such as the one presented here, allows to test the benefits of the increased model complexity. Such extension of the OGGM are obviously not limited to simple model, but may as well add more complex processes \citep[e.g., an improved calving parametrization,][]{Recinos2019}. It shows the benefits of an open source, modular, community driven glacier model. Comparing different parametrization and estimating uncertainties are imperative to improve projections of future glacier mass loss and its consequences. The global implementation of the scaling model and the implementation of additional geometry parameterizations \citep[e.g., GloGEM by][]{Huss2015} could give further insights into the benefits and limitations of explicit glacier dynamics.
