%!TEX root = thesis.tex
\newcommand{\advgeophys}{Adv. Geophys} 
\newcommand{\agrformet}{Agric. and For. Meteorol.} 
\newcommand{\agrmet}{Agric. Meteorol.} 
\newcommand{\anngeophysicae}{Ann. Geophysicae} 
\newcommand{\annhy}{Ann. Hydrogr} 
\newcommand{\annmetnf}{Ann. Metorol., N.F.} 
\newcommand{\annmet}{Ann. Metorol} 
\newcommand{\annrevfluidmech}{Ann. Rev. Fluid Mech.} 
\newcommand{\archmet}{Arch. Meteorol} 
\newcommand{\archmgb}{Arch. Met. Geoph. Biokl} 
\newcommand{\atmosres}{Atmos. Res.} 
\newcommand{\bams}{Bull. Amer. Meteorol. Soc.} 
\newcommand{\beitrphysatmos}{Beitr. Phys. Atmos.} 
\newcommand{\blmet}{Bound.-Layer. Meteor.} 
\newcommand{\bonnmet}{Bonner Met. Abh.}  
\newcommand{\bulloegm}{Bull. \"Oster. Gesellsch. f\"ur Meteorologie}
\newcommand{\climatedyn}{Climate Dyn.} 
\newcommand{\dynatmos}{Dyn. Atmos.} 
\newcommand{\dynatoc}{Dyn. Atmos. Oceans} 
\newcommand{\flumech}{Fluid Mechanics} 
\newcommand{\gafd}{Geophys. Astrophys. Fluid Dynamics} 
\newcommand{\geophysfludyn}{Geophysical Fluid Dynamics} 
\newcommand{\geophysica}{Geophysica} 
\newcommand{\geophysmono}{Geophys. Monogr.} 
\newcommand{\geordsch}{Geogr. Rdsch.} 
\newcommand{\globpc}{Global Planet. Change}
\newcommand{\igeophysser}{Int. Geophys. Ser} 
\newcommand{\ijclimogy}{Int. J. Climatol.} 
\newcommand{\ijremote}{Int. J. Remote Sensing} 
\newcommand{\issatocmod}{Iss. Atmos. Ocean Mod.}
\newcommand{\jamstatasso}{J. Amer. Stat. Assoc.} 
\newcommand{\japplecol}{J. Appl. Ecol} 
\newcommand{\japplmet}{J. Appl. Meteor.} 
\newcommand{\jas}{J. Atmos. Sci.} 
\newcommand{\jclimapplmet}{J. Clim. Appl. Meteor} 
\newcommand{\jclimatology}{J. Climatology} 
\newcommand{\jcomplex}{J. Complexity} 
\newcommand{\jcompphys}{J. Comp. Phys.} 
\newcommand{\jc}{J. Climate} 
\newcommand{\jdemathpea}{J. de Math\'ematique pures et appliqu} 
\newcommand{\jflumech}{J. Fluid Mech.} 
\newcommand{\jgr}{J. Geophys. Res.} 
\newcommand{\jmarres}{J. Mar. Res.} 
\newcommand{\jmetjap}{J. Meteor. Soc. Jap.} 
\newcommand{\jmetorol}{J. Meteorol.} 
\newcommand{\jpo}{J. Phys. Oceanogr.} 
\newcommand{\jresnbs}{J. Res. Nat. Bureau Standards} 
\newcommand{\mathannal}{Math. Ann.}  
\newcommand{\mathcomp}{Math. Comp.} 
\newcommand{\map}{Meteorol. Atmos. Phys.} 
\newcommand{\metatmphy}{Meteorol. Atmos. Phys.} 
\newcommand{\metrdsch}{Meteorol. Rdsch.}  
\newcommand{\mittdmg}{Mitt. Dtsch. Meteorol. Gesellsch.} 
\newcommand{\mwr}{Mon. Wea. Rev.} 
\newcommand{\mznf}{Meteorol. Z., N.F.} 
\newcommand{\mz}{Meteorol.Z.} 
\newcommand{\philtransroylondon}{Philos. Trans. Roy. Soc. London} 
\newcommand{\physclim}{Phys. Clim.} 
\newcommand{\physflud}{Phys. Fluids} 
\newcommand{\physrev}{Phys. Rev.} 
\newcommand{\qjrms}{Quart. J. Roy. Meteor. Soc.} 
\newcommand{\revgeophys}{Rev. Geophys} 
\newcommand{\revgeospac}{Rev. Geoph. Space. Phys.} 
\newcommand{\revmodphys}{Rev. Modern Phys.} 
\newcommand{\science}{Science} 
\newcommand{\simarev}{SIAM Rev.} 
\newcommand{\tellus}{Tellus} 
\newcommand{\weaforc}{Wea. Forecasting} 
\newcommand{\wetterleben}{Wetter und Leben} 
